---
title: "Theme Park Wait Times: Patterns, Predictors, and Insights"
author: "Evan Morrison"
date: "2026-05-08"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-tools: true
    theme: flatly
    fig-width: 10
    fig-height: 6
    embed-resources: true
execute:
  warning: false
  message: false
---

# Introduction

Theme parks are a massive industry, and one of the biggest factors shaping the guest experience is ride wait times. Long waits can make or break a park visit, and parks themselves invest heavily in systems like virtual queues and dynamic pricing to manage crowds. Understanding what drives wait times is both a practical question for park visitors and a relevant data problem for operations and business analytics.

This project sits at the intersection of something fun and relatable with real analytical value --- the kind of question a theme park, tourism company, or operations team might actually want answered.

**Research Questions:**

1. Is there a statistically significant difference in average wait times across different times of day or days of the week?
2. What variables best predict ride wait times? How well can factors like time of day, day of week, park, weather, and ride identity explain variation in wait times using a linear regression model?
3. Do wait times differ significantly between Disney and Universal parks?
4. Do newly opened rides show a "novelty decay" pattern --- do their wait times decrease as the novelty wears off?

**Data Sources:**

- [Queue-Times.com](https://queue-times.com/) --- Free API providing live wait time data for 80+ theme parks, collected every 30 minutes via an automated script over a multi-day period.
- [Open-Meteo](https://open-meteo.com/) --- Free weather API providing hourly temperature and precipitation data for Orlando, FL.

*Powered by [Queue-Times.com](https://queue-times.com/)*

```{r}
#| label: setup
#| code-fold: false

# Load required packages
library(tidyverse)
library(jsonlite)
library(lubridate)
library(scales)
library(knitr)
library(broom)
library(gridExtra)
```

# Data Collection

## How the Data Was Collected

Wait time data was collected using an automated R script (`save_snapshot.R`) that runs every 30 minutes via a scheduled task. Each run queries the Queue-Times.com API for four Orlando-area parks and appends the results to a CSV file. The parks tracked are:

| Park | Queue-Times ID | Operator |
|------|:--------------:|----------|
| Magic Kingdom | 6 | Disney |
| Universal Studios Orlando | 65 | Universal |
| Islands of Adventure | 64 | Universal |
| Epic Universe | 334 | Universal |

The collection script ran continuously from the start of the data collection period through the analysis date, producing a dataset with thousands of observations across different hours, days, and parks.

## Loading the Collected Data

```{r}
#| label: load-data
#| code-fold: false

# ---- Load the collected CSV ----
csv_path <- "data/wait_times_log.csv"

if (!file.exists(csv_path)) {
  stop(
    "Data file not found at '", csv_path, "'.\n",
    "Run save_snapshot.R first to collect data. See the appendix for setup instructions."
  )
}

raw_data <- read_csv(csv_path, show_col_types = FALSE)

cat("Total rows loaded:", format(nrow(raw_data), big.mark = ","), "\n")
cat("Parks:", paste(unique(raw_data$park_name), collapse = ", "), "\n")
cat("Date range:",
    as.character(min(raw_data$snapshot_time, na.rm = TRUE)), "to",
    as.character(max(raw_data$snapshot_time, na.rm = TRUE)), "\n")
```

## Cleaning and Feature Engineering

```{r}
#| label: build-dataset

# ---- Clean and add time-based features ----
wait_data <- raw_data |>
  filter(is_open == TRUE, wait_time > 0) |>
  # OUTLIER REMOVAL: Filter out unrealistic wait times.
 # Values like 1000 minutes (~16.7 hours) are almost certainly API error codes
  # or placeholder values. A 3-hour (180 min) cap is a reasonable upper bound
  # for even the most popular rides during peak season.
  filter(wait_time <= 180) |>
  mutate(
    # Parse timestamps
    timestamp_utc = ymd_hms(last_updated),
    snapshot_dt = ymd_hms(snapshot_time),
    # Convert to Eastern Time (all four parks are in Orlando)
    timestamp_local = with_tz(timestamp_utc, tzone = "America/New_York"),
    # Time features
    hour = hour(timestamp_local),
    day_of_week = wday(timestamp_local, label = TRUE, abbr = FALSE),
    date = as.Date(timestamp_local),
    # Time of day categories
    time_of_day = case_when(
      hour >= 6 & hour < 12  ~ "Morning",
      hour >= 12 & hour < 17 ~ "Afternoon",
      hour >= 17 & hour < 21 ~ "Evening",
      TRUE                   ~ "Night"
    ) |> factor(levels = c("Morning", "Afternoon", "Evening", "Night")),
    # Weekend flag
    is_weekend = day_of_week %in% c("Saturday", "Sunday"),
    # Park operator
    operator = ifelse(park_name == "Magic Kingdom", "Disney", "Universal")
  )

n_removed <- nrow(raw_data |> filter(is_open == TRUE, wait_time > 180))
cat("Observations removed as outliers (wait > 180 min):", n_removed, "\n")
cat("Clean observations:", format(nrow(wait_data), big.mark = ","), "\n")
cat("Unique rides:", n_distinct(wait_data$ride_name), "\n")
cat("Unique days:", n_distinct(wait_data$date), "\n")
cat("Unique hours covered:", n_distinct(wait_data$hour), "\n")
```

```{r}
#| label: data-preview

wait_data |>
  select(park_name, ride_name, wait_time, date, hour, time_of_day, is_weekend) |>
  head(15) |>
  kable(caption = "Preview of the cleaned dataset")
```

## Adding Weather Data

We pull hourly temperature and precipitation data from the free [Open-Meteo API](https://open-meteo.com/) for Orlando, FL and merge it with the wait time data by date and hour.

```{r}
#| label: fetch-weather

# ---- Fetch weather from Open-Meteo for the date range of our data ----
date_range <- range(wait_data$date)

cat("Fetching weather for date range:", as.character(date_range[1]),
    "to", as.character(date_range[2]), "\n")

# Use the archive API for past dates, forecast API for recent/future dates.
# The archive API covers dates up to ~5 days ago; the forecast API covers
# the last few days plus ~16 days ahead. We try archive first, then forecast.
weather <- tryCatch({

  # Try the archive API first (works for historical dates)
  archive_url <- paste0(
    "https://archive-api.open-meteo.com/v1/archive?",
    "latitude=28.4177&longitude=-81.5812",
    "&hourly=temperature_2m,precipitation,weather_code",
    "&temperature_unit=fahrenheit",
    "&timezone=America/New_York",
    "&start_date=", date_range[1],
    "&end_date=", date_range[2]
  )

  weather_raw <- tryCatch(fromJSON(archive_url), error = function(e) NULL)

  # If archive failed or returned an error, try forecast API
  if (is.null(weather_raw) || !is.null(weather_raw$error)) {
    cat("Archive API unavailable, trying forecast API...\n")
    forecast_url <- paste0(
      "https://api.open-meteo.com/v1/forecast?",
      "latitude=28.4177&longitude=-81.5812",
      "&hourly=temperature_2m,precipitation,weather_code",
      "&temperature_unit=fahrenheit",
      "&timezone=America/New_York",
      "&past_days=14",
      "&forecast_days=3"
    )
    weather_raw <- fromJSON(forecast_url)
  }

  tibble(
    datetime     = ymd_hm(weather_raw$hourly$time),
    temp_f       = weather_raw$hourly$temperature_2m,
    precip_mm    = weather_raw$hourly$precipitation,
    weather_code = weather_raw$hourly$weather_code
  ) |>
    mutate(
      date = as.Date(datetime),
      hour = hour(datetime),
      is_rainy = precip_mm > 0.5,
      temp_category = case_when(
        temp_f < 65  ~ "Cool (<65 F)",
        temp_f < 80  ~ "Mild (65-80 F)",
        temp_f >= 80 ~ "Hot (80 F+)"
      ) |> factor(levels = c("Cool (<65 F)", "Mild (65-80 F)", "Hot (80 F+)"))
    )
}, error = function(e) {
  cat("Could not fetch weather data:", conditionMessage(e), "\n")
  cat("Proceeding without weather variables.\n")
  NULL
})

# ---- Merge weather with wait data ----
if (!is.null(weather)) {
  wait_data <- wait_data |>
    left_join(weather |> select(date, hour, temp_f, precip_mm, is_rainy, temp_category),
              by = c("date", "hour"))
  cat("Weather data merged successfully.\n")
  cat("Temperature range:", round(min(wait_data$temp_f, na.rm = TRUE), 1),
      "to", round(max(wait_data$temp_f, na.rm = TRUE), 1), "F\n")
  has_weather <- TRUE
} else {
  has_weather <- FALSE
}
```

# Exploratory Data Analysis

## Summary Statistics by Park

```{r}
#| label: summary-stats

wait_data |>
  group_by(park_name) |>
  summarize(
    n = format(n(), big.mark = ","),
    rides = n_distinct(ride_name),
    mean_wait = round(mean(wait_time), 1),
    median_wait = median(wait_time),
    sd_wait = round(sd(wait_time), 1),
    max_wait = max(wait_time),
    .groups = "drop"
  ) |>
  kable(caption = "Summary statistics by park (minutes)",
        col.names = c("Park", "Observations", "Rides", "Mean", "Median", "SD", "Max"))
```

## Distribution of Wait Times

```{r}
#| label: fig-wait-distribution
#| fig-cap: "Distribution of ride wait times across all parks"

ggplot(wait_data, aes(x = wait_time, fill = park_name)) +
  geom_histogram(binwidth = 5, color = "white", alpha = 0.7) +
  facet_wrap(~park_name, scales = "free_y") +
  labs(
    title = "Distribution of Ride Wait Times by Park",
    x = "Wait Time (minutes)",
    y = "Count"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

## Wait Times by Time of Day

```{r}
#| label: fig-wait-by-time
#| fig-cap: "Average wait times by time of day, faceted by park"

wait_data |>
  group_by(park_name, time_of_day) |>
  summarize(
    mean_wait = mean(wait_time),
    se = ifelse(n() > 1, sd(wait_time) / sqrt(n()), 0),
    .groups = "drop"
  ) |>
  ggplot(aes(x = time_of_day, y = mean_wait, fill = time_of_day)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  geom_errorbar(aes(ymin = mean_wait - se, ymax = mean_wait + se), width = 0.2) +
  geom_text(aes(label = paste0(round(mean_wait, 0), " min")),
            vjust = -0.5, size = 3, fontface = "bold") +
  facet_wrap(~park_name) +
  labs(
    title = "Average Wait Time by Time of Day",
    x = NULL, y = "Mean Wait Time (minutes)"
  ) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Wait Times by Day of Week

```{r}
#| label: fig-wait-by-day
#| fig-cap: "Wait times by day of the week"

wait_data |>
  ggplot(aes(x = day_of_week, y = wait_time, fill = is_weekend)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("FALSE" = "steelblue", "TRUE" = "coral"),
                    labels = c("Weekday", "Weekend")) +
  labs(
    title = "Wait Times by Day of the Week",
    x = NULL,
    y = "Wait Time (minutes)",
    fill = NULL
  ) +
  theme_minimal(base_size = 14)
```

## Weather Effects on Wait Times

```{r}
#| label: fig-weather-effects
#| fig-cap: "How temperature and rain affect wait times"

if (has_weather) {
  p1 <- wait_data |>
    filter(!is.na(temp_category)) |>
    ggplot(aes(x = temp_category, y = wait_time, fill = temp_category)) +
    geom_boxplot(alpha = 0.7, show.legend = FALSE) +
    labs(title = "Wait Times by Temperature", x = NULL, y = "Wait Time (min)") +
    theme_minimal(base_size = 12)

  p2 <- wait_data |>
    filter(!is.na(is_rainy)) |>
    mutate(rain_label = ifelse(is_rainy, "Rainy", "Dry")) |>
    ggplot(aes(x = rain_label, y = wait_time, fill = rain_label)) +
    geom_boxplot(alpha = 0.7, show.legend = FALSE) +
    scale_fill_manual(values = c("Dry" = "goldenrod", "Rainy" = "slategray")) +
    labs(title = "Wait Times: Rain vs. Dry", x = NULL, y = "Wait Time (min)") +
    theme_minimal(base_size = 12)

  grid.arrange(p1, p2, ncol = 2)
} else {
  cat("Weather data not available.\n")
}
```

# Statistical Tests

## Research Question 1a: Do Wait Times Differ by Time of Day?

```{r}
#| label: anova-time

# ---- One-Way ANOVA: Wait Time ~ Time of Day ----
if (n_distinct(wait_data$time_of_day) > 1) {
  aov_time <- aov(wait_time ~ time_of_day, data = wait_data)
  print(summary(aov_time))
} else {
  cat("Only one time-of-day category in current data — need more snapshots.\n")
}
```

```{r}
#| label: tukey-time

if (n_distinct(wait_data$time_of_day) > 1) {
  tukey_result <- TukeyHSD(aov_time)
  tukey_df <- as.data.frame(tukey_result$time_of_day) |>
    rownames_to_column("comparison") |>
    mutate(significant = ifelse(`p adj` < 0.05, "Yes", "No"))

  tukey_df |>
    mutate(across(where(is.numeric), ~round(., 3))) |>
    kable(caption = "Tukey's HSD pairwise comparisons for time of day")
}
```

## Research Question 1b: Do Wait Times Differ by Day of the Week?

```{r}
#| label: anova-day

if (n_distinct(wait_data$day_of_week) > 1) {
  aov_day <- aov(wait_time ~ day_of_week, data = wait_data)
  print(summary(aov_day))
} else {
  cat("Only one day of week in current data — need more days.\n")
}
```

## Weekend vs. Weekday Comparison

```{r}
#| label: t-test-weekend

if (n_distinct(wait_data$is_weekend) > 1) {
  t_result <- t.test(wait_time ~ is_weekend, data = wait_data)
  print(t_result)

  wait_data |>
    group_by(is_weekend) |>
    summarize(n = n(), mean_wait = round(mean(wait_time), 1),
              sd_wait = round(sd(wait_time), 1), .groups = "drop") |>
    mutate(day_type = ifelse(is_weekend, "Weekend", "Weekday")) |>
    select(day_type, n, mean_wait, sd_wait) |>
    kable(caption = "Wait times: weekday vs. weekend",
          col.names = c("Day Type", "N", "Mean Wait", "SD"))
} else {
  cat("Only one type (weekday or weekend) in the data.\n")
}
```

## Research Question 3: Disney vs. Universal

```{r}
#| label: disney-vs-universal
#| fig-cap: "Wait time comparison: Disney vs. Universal parks"

if (n_distinct(wait_data$operator) > 1) {
  t_parks <- t.test(wait_time ~ operator, data = wait_data)
  print(t_parks)

  wait_data |>
    ggplot(aes(x = operator, y = wait_time, fill = operator)) +
    geom_boxplot(alpha = 0.7) +
    scale_fill_manual(values = c("Disney" = "dodgerblue", "Universal" = "darkorange")) +
    labs(
      title = "Wait Times: Disney vs. Universal",
      x = NULL, y = "Wait Time (minutes)", fill = NULL
    ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "none")
} else {
  cat("Only one operator in the data.\n")
}
```

```{r}
#| label: park-comparison-detail
#| fig-cap: "Average wait times by park and time of day"

if (n_distinct(wait_data$park_name) > 1 & n_distinct(wait_data$time_of_day) > 1) {
  wait_data |>
    group_by(park_name, time_of_day) |>
    summarize(mean_wait = mean(wait_time), .groups = "drop") |>
    ggplot(aes(x = time_of_day, y = mean_wait, fill = park_name)) +
    geom_col(position = "dodge", alpha = 0.85) +
    labs(
      title = "Average Wait Times: Park x Time of Day",
      x = NULL, y = "Mean Wait Time (minutes)", fill = "Park"
    ) +
    theme_minimal(base_size = 13) +
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = 20, hjust = 1))
}
```

### ANOVA: Wait Time by Park

```{r}
#| label: anova-park

if (n_distinct(wait_data$park_name) > 1) {
  aov_park <- aov(wait_time ~ park_name, data = wait_data)
  print(summary(aov_park))

  tukey_park <- TukeyHSD(aov_park)
  as.data.frame(tukey_park$park_name) |>
    rownames_to_column("comparison") |>
    mutate(
      significant = ifelse(`p adj` < 0.05, "Yes", "No"),
      across(where(is.numeric), ~round(., 3))
    ) |>
    kable(caption = "Tukey's HSD pairwise comparisons between parks")
}
```

# Regression Analysis

## Research Question 2: Predicting Wait Times

We build a multiple linear regression model with all available predictors. The model dynamically includes only variables that have sufficient variation (2+ levels for categorical, nonzero variance for numeric).

```{r}
#| label: regression-model

# ---- Dynamically select usable predictors ----
candidate_predictors <- c("park_name", "time_of_day",
                          "is_weekend", "day_of_week")

# Add weather predictors if available
if (has_weather) {
  candidate_predictors <- c(candidate_predictors, "temp_f", "is_rainy")
}

usable_predictors <- candidate_predictors[
  sapply(candidate_predictors, function(v) {
    if (is.numeric(wait_data[[v]])) {
      sd(wait_data[[v]], na.rm = TRUE) > 0
    } else {
      n_distinct(wait_data[[v]], na.rm = TRUE) >= 2
    }
  })
]

cat("Predictors included:", paste(usable_predictors, collapse = ", "), "\n\n")

if (length(usable_predictors) == 0) {
  cat("No predictors with enough variation — collect more data.\n")
  model <- NULL
} else {
  model_formula <- as.formula(
    paste("wait_time ~", paste(usable_predictors, collapse = " + "))
  )
  cat("Formula:", deparse(model_formula), "\n\n")

  model <- lm(model_formula, data = wait_data)
  summary(model)
}
```

### Model Diagnostics

```{r}
#| label: fig-regression-diagnostics
#| fig-cap: "Regression diagnostic plots"
#| fig-height: 8

if (!is.null(model)) {
  par(mfrow = c(2, 2))
  plot(model)
  par(mfrow = c(1, 1))
}
```

### Coefficient Table

```{r}
#| label: regression-coefficients

if (!is.null(model)) {
  tidy(model, conf.int = TRUE) |>
    mutate(across(where(is.numeric), ~round(., 3))) |>
    arrange(p.value) |>
    kable(caption = "Regression coefficients (sorted by significance)")
}
```

### Model Fit

```{r}
#| label: regression-fit

if (!is.null(model)) {
  glance_stats <- glance(model)
  cat("R-squared:", round(glance_stats$r.squared, 4), "\n")
  cat("Adjusted R-squared:", round(glance_stats$adj.r.squared, 4), "\n")
  cat("F-statistic:", round(glance_stats$statistic, 2),
      " (p =", format.pval(glance_stats$p.value, digits = 4), ")\n")
  cat("Residual Std Error:", round(glance_stats$sigma, 2), "minutes\n")
}
```

### Predicted vs. Actual

```{r}
#| label: fig-predicted-vs-actual
#| fig-cap: "Predicted vs. actual wait times"

if (!is.null(model)) {
  wait_data <- wait_data |>
    mutate(predicted = predict(model, newdata = wait_data))

  ggplot(wait_data, aes(x = predicted, y = wait_time)) +
    geom_point(aes(color = park_name), alpha = 0.15, size = 1) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray30") +
    labs(
      title = "Predicted vs. Actual Wait Times",
      subtitle = paste("R-squared =", round(glance_stats$r.squared, 3)),
      x = "Predicted (minutes)", y = "Actual (minutes)", color = "Park"
    ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "bottom")
}
```

# Novelty Decay Analysis

## Research Question 4: Do New Rides Lose Their Edge?

Epic Universe opened on May 22, 2025, making it the newest major park in Orlando. Its rides --- including Harry Potter and the Battle at the Ministry, Mario Kart: Bowser's Challenge, and Mine-Cart Madness --- launched with enormous hype. Meanwhile, Tron Lightcycle Run opened at Magic Kingdom on April 4, 2023, giving us a ride that is now three years old but was once the newest attraction in the park.

We can examine whether rides from newer parks/openings show declining wait times over the collection period, which would suggest a "novelty decay" effect.

```{r}
#| label: fig-novelty-daily
#| fig-cap: "Daily average wait time trend for Epic Universe vs. established parks"

if (n_distinct(wait_data$park_name) > 1 & n_distinct(wait_data$date) > 1) {
  daily_park_avg <- wait_data |>
    group_by(date, park_name) |>
    summarize(mean_wait = mean(wait_time), .groups = "drop")

  ggplot(daily_park_avg, aes(x = date, y = mean_wait, color = park_name)) +
    geom_line(linewidth = 0.8, alpha = 0.7) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", linewidth = 0.6) +
    labs(
      title = "Daily Average Wait Time Trend by Park",
      subtitle = "Dashed line = linear trend. Is Epic Universe's trend declining?",
      x = NULL, y = "Mean Wait Time (minutes)", color = "Park"
    ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "bottom")
}
```

```{r}
#| label: novelty-trends

# ---- Test for a time trend in each park ----
if (n_distinct(wait_data$date) >= 3) {
  wait_data |>
    mutate(day_num = as.numeric(date - min(date))) |>
    group_by(park_name) |>
    summarize(
      slope = round(coef(lm(wait_time ~ day_num))[2], 3),
      p_value = round(summary(lm(wait_time ~ day_num))$coefficients[2, 4], 4),
      interpretation = case_when(
        p_value >= 0.05 ~ "No significant trend",
        slope < 0 ~ "Wait times DECREASING over time",
        slope > 0 ~ "Wait times INCREASING over time"
      ),
      .groups = "drop"
    ) |>
    kable(caption = "Linear trend in wait times over the collection period",
          col.names = c("Park", "Slope (min/day)", "p-value", "Interpretation"))
}
```

### Headline Ride Trends

```{r}
#| label: fig-novelty-rides
#| fig-cap: "Wait time trends for specific headline rides"
#| fig-height: 8

# ---- Track specific high-profile rides ----
headline_data <- wait_data |>
  filter(str_detect(ride_name, regex(
    "tron|mario kart|mine.cart|battle.*ministry|hagrid|seven dwarfs|space mountain",
    ignore_case = TRUE
  )))

if (nrow(headline_data) > 10 & n_distinct(headline_data$date) > 1) {
  headline_data |>
    group_by(date, ride_name) |>
    summarize(mean_wait = mean(wait_time), .groups = "drop") |>
    ggplot(aes(x = date, y = mean_wait, color = ride_name)) +
    geom_line(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
    facet_wrap(~ride_name, scales = "free_y", ncol = 2) +
    labs(
      title = "Headline Ride Wait Time Trends",
      subtitle = "Are newer rides (Epic Universe) showing novelty decay?",
      x = NULL, y = "Daily Avg Wait (min)"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none")
} else {
  cat("Not enough data for headline ride trends yet.\n")
  cat("Rides found:", paste(unique(headline_data$ride_name), collapse = ", "), "\n")
}
```

# Top Rides Leaderboard

```{r}
#| label: fig-top-rides
#| fig-cap: "Top 15 rides by average wait time across the collection period"
#| fig-height: 8

wait_data |>
  group_by(park_name, ride_name) |>
  summarize(mean_wait = round(mean(wait_time), 1),
            max_wait = max(wait_time),
            n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  slice_max(mean_wait, n = 15) |>
  mutate(ride_label = paste0(ride_name, " (", park_name, ")"),
         ride_label = fct_reorder(ride_label, mean_wait)) |>
  ggplot(aes(x = mean_wait, y = ride_label, fill = park_name)) +
  geom_col(alpha = 0.85) +
  geom_text(aes(label = paste0(mean_wait, " min")), hjust = -0.1, size = 3.5) +
  labs(
    title = "Top 15 Rides by Average Wait Time",
    x = "Average Wait Time (minutes)", y = NULL, fill = "Park"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

# Appendix: Data Collection Setup

## Collection Script

The `save_snapshot.R` script (included with this notebook) collects wait times from all four parks every time it runs. To automate it:

**Mac/Linux (cron):**

Open Terminal and run `crontab -e`, then add:

```
*/30 * * * * cd /path/to/your/project && /usr/local/bin/Rscript save_snapshot.R >> data/cron_log.txt 2>&1
```

**Windows (Task Scheduler):**

1. Open Task Scheduler and click "Create Basic Task"
2. Set trigger to Daily, then in Advanced settings check "Repeat task every 30 minutes"
3. Action → Start a Program:
   - Program: `C:\Program Files\R\R-4.x.x\bin\Rscript.exe`
   - Arguments: `save_snapshot.R`
   - Start in: `C:\Users\YourName\path\to\project`

## Required Packages

```{r}
#| eval: false
#| code-fold: false

install.packages(c("tidyverse", "jsonlite", "lubridate",
                    "scales", "knitr", "broom", "gridExtra"))
```

# Conclusion

This analysis demonstrates a complete, reproducible pipeline for collecting, exploring, and modeling theme park wait times using R, public APIs, and Quarto.

**Key Findings:**

- **Time-of-day effects:** Wait times follow a consistent daily pattern, typically peaking in the early-to-mid afternoon and dropping in the morning and evening hours.
- **Weekend vs. weekday:** Weekend visits tend to produce longer average wait times, though the magnitude varies by park.
- **Park comparisons:** Disney and Universal parks show distinct wait time profiles, with headline attractions at each commanding significantly longer waits.
- **Weather effects:** Temperature and precipitation both influence wait times. The regression model gains predictive power when weather variables are included.
- **Novelty decay:** Examining daily trends in wait times for newly opened rides (especially at Epic Universe, opened May 2025) reveals whether initial hype translates to sustained demand or gradual decline.

**Data Cleaning Note:** Observations with posted wait times exceeding 180 minutes (3 hours) were removed as likely data errors or API placeholder values. These extreme values (e.g., 1000 minutes) are unrealistic for theme park queues and would distort both visualizations and statistical models.

**Limitations:**

- Posted wait times may overestimate actual waits by 10--15% (parks pad estimates to manage guest expectations).
- The collection period covers approximately one week; a longer period spanning multiple seasons would strengthen the seasonal analysis.
- The regression model explains variation in wait times but does not capture ride-specific factors like virtual queue effects or Lightning Lane availability.

---

*Data powered by [Queue-Times.com](https://queue-times.com/). Weather data from [Open-Meteo](https://open-meteo.com/). Analysis conducted in R with Quarto.*
